Speculation in foreign exchange: noise or information

### 2015-06-02 Knitr/NewPaper.Rnw
Work and Home aligned. Literature review progressing.  Please enter the commit message for your changes. Lines starting
Changes on home pc.  Everything seems to be fine. Now completing the last of the lit rev.

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
%\bibliographystyle{plainnat}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
 \usepackage{listings}
  \usepackage{tikz}
 \usetikzlibrary{arrows,positioning}
 \usepackage{color}
 %\graphicspath{{../Pictures/}}
\definecolor{mygrey}{gray}{0.95}
\lstset{backgroundcolor=\color{mygrey}}
\begin{document}
\title{Speculation in foreign exchange: noise or information}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk. }} 
\date{\today}
\maketitle
\begin{abstract}
An event study of extremes is used to assess the Keynesian and Friedman views of speculation.    Speculative extremes are ranked by intensity of sentiment and weight of activity.  The extreme says little about the future in either case, suggesting that speculation is more than just random noise and supporting the view there is some informational content that is being passed through to the price with speculative activity. A unique dataset of risk-reversal skew on option prices is used to measure the intensity of speculative sentiment.  This is supplemented with regulatory data on the weight of speculative positions.  
\end{abstract}
\section*{Introduction}
There are a range of ideas about the way that speculation can affect financial markets.  Towards one end of this spectrum is the view of Keynes from Chapter 13 of \emph{The General Theory} where he identified speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citet[p. 101]{Keynes1936}. Here speculative activity is a problem that, at its most extreme, will cause booms, reversals and will discourage fundamental, long-term investors from participating. It is noise that does not emerge randomly, but feeds off itself as a collective sentiment  encourages  trends and serial correlation in asset prices.

Towards the other end, Milton Friedman argued that speculation was a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. In arguing in favour of flexible exchange rates, Friedman asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price \citet{FriedmanPositive}. Therefore, Friedmand argued, the speculative process would favour those able to access or utilise information over the uninformed and, through a Darwinian process, informed speculation would dominate.  This view regards speculation as an key lubricant of financial market as it provides liquidity and facilitates price discovery.

The aim of this paper is to understand more about the nature of speculation by investigating the relationship between both the intensity of speculation and the weight of speculative positions with the movement of prices in the foreign exchange market.   If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, and when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.

A unique set of option prices are used to identify the intensity of speculative sentiment.  US regulatory positions measure the prominence of speculators in the market. Speculation does not appear to be random, there is momentum to the rise and fall in sentiment and positions.  However, while currency prices co-move with sentiment and positions, there is no informational content from these positions, meaning that knowing the positions does not imoprove forecasts of future returns, supporting the suggestion that speculative activity has a permanent affect on prices and that is consistent with speculation being part of the process of price discovery. Speculators do not seem to drive prices from fundamental value and their presence does not seem to hamper the incorporation of information into prices.

@@

In the \emph{microstructure} literature, the term noise-trader is interchanged with \emph{liquidity-trader} to mean those trading with a lack of information and those trading for liquidity reasons.  They do not usually distinguish between psychological and institutional reasons for the lack of attention to fundamental information. Nonetheless, liquidity tends to refer to institutional explanations for trading without reference to fundamental value: these agents trade without reference to value.  This may mean that they just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.  Examples of the microstructure approach are found in \citet{Kyle1985Continuous} and \citet{glosten1985bid}.

The other issue is whether uninformed speculative noise-trading, of either variety, can be considered as just random noise?   The original microstrucutre version had moise-traders arriving in a random fashion so that the market-maker could infer information from an imbalance of orders \citet{Kyle1985Continuous}.  However, though the idea that noise is a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, this exogenous construct may be an over-simplification that is not apprpropriate when looking for more understanding of speculation itself.

@@

 Shliefer, Summers, DeLong etc identify the limits to arbitrage. The other writings are Battalio, Hatch and Jennings (1997), Harris and Schultz (1998) and Scheinkman and Xiong (2003).  Lowenstein and Willard (2006)

It would be good to have some more modern ideas here.  Maybe they are just the empirical results. 
\subsection{Interplay between Informed and uninformed}
\citet[p. 1]{GortonNoise}  say that noise-traders are all those trading in markets for ``non-information-based reasons'', and that this was originally a solution to the problem of understanding or justifying the trading of securities: if economic agents know the fundamental value of a security how can some be buying and some be selling.  While will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of it. Informed traders, knowing that the current price is below fundamental value, can purchase from noise-traders, who either do not know or do not care.  Noise-traders can therefore create a gap between the price and the fundamtental value that will encourage informed traders to pay the cost of acquiring information. The divergence between this price and fundametnal value is essential to encourage te process of price discovery.

@@

  Why do noise traders exist?  There are liquidity and behavioural explanations. There are also a number of models that show noise traders as momentum traders.

@@

One approach to this problem is to use the evidence that noise-traders utilise indicators or techniques that are believed to provide information when they do not.  \href{http://papers.ssrn.com.ezproxy.liv.ac.uk/sol3/papers.cfm?abstract_id=869827}{do noise traders move markets?}  there is strong evidence of herding by individual investors. there is also momentum. there is reversal:  stocks that were heavily bought one year under perform; stocks that were heavily sold, outperform. the opposite in the short run:  heavily bought one week generate large returns the next week.  \citet{ShillerFashion} discusses fashionable techniques.  In this way, speculation noise is associated with sentiment and measures of sentiment are then used as ways of measuring speculation. In addition,empirical work will often identify speculation or noise with momentum trading.

>>>>

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
%\bibliographystyle{plainnat}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
 \usepackage{listings}
  \usepackage{tikz}
 \usetikzlibrary{arrows,positioning}
 \usepackage{color}
 %\graphicspath{{../Pictures/}}
\definecolor{mygrey}{gray}{0.95}
\lstset{backgroundcolor=\color{mygrey}}
\begin{document}
\title{Speculation in foreign exchange: noise or information}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk. }} 
\date{\today}
\maketitle
\begin{abstract}
An event study is used to assess the Keynesian and Friedman views of speculation.    Speculative extremes are ranked by intensity of sentiment and weight of activity.  The extreme says little about the future in either case, suggesting that speculation is more than just random noise and supporting the view there is some informational content that is being passed through to the price with speculative activity. A unique dataset of risk-reversal skew on option prices is used to measure the intensity of speculative sentiment; the weight of positions is based on US regulatory data.  
\end{abstract}
\section*{Introduction}
There are a range of ideas about the way that speculation can affect financial markets.  Towards one end of this spectrum is the view of Keynes from Chapter 13 of \emph{The General Theory} where he identified speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citet[p. 101]{Keynes1936}. Here speculative activity is a problem that, at its most extreme, will cause booms, reversals and will discourage fundamental, long-term investment. It is noise that does not emerge randomly, but feeds off itself as a collective sentiment  encourages  trends and serial correlation in asset prices.

Towards the other end, Milton Friedman argued that speculation was a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. In arguing in favour of flexible exchange rates, Friedman asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price \citet{FriedmanPositive}. Therefore, Friedmand argued, the speculative process would favour those able to access or utilise information over the uninformed and, through a Darwinian process, informed speculation would dominate.  This view regards speculation as an key lubricant of financial market as it provides liquidity and facilitates the process of price discovery.

The aim of this paper is to understand more about the nature of speculation by investigating the relationship between both the intensity of speculation and the weight of speculative positions with the movement of prices in the foreign exchange market.   If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.

A unique set of option prices are used to identify the intensity of speculative sentiment.  US regulatory positions measure the prominence of speculators in the market. Speculation does not appear to be random, there is momentum to the rise and fall in sentiment and positions.  However, while currency prices co-move with sentiment and positions, there is no informational content from either. As such, knowing the positions does not improve forecasts of future returns. This supports the suggestion that speculative activity has a permanent affect on prices and that is consistent with speculation being part of the process of price discovery. Speculators do not seem to drive prices from fundamental value and their presence does not seem to hamper the incorporation of information into prices.

@@

In the \emph{microstructure} literature, the term noise-trader is interchanged with \emph{liquidity-trader} to mean those trading with a lack of information and those trading for liquidity reasons.  They do not usually distinguish between psychological and institutional reasons for the lack of attention to fundamental information. Nonetheless, liquidity tends to refer to institutional explanations for trading without reference to fundamental value. This may mean that they just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.  Examples of the microstructure approach are found in \citet{Kyle1985Continuous} and \citet{glosten1985bid}.

The other issue is whether uninformed speculative noise-trading, of either variety, can be considered as just random noise?   The original microstrucutre version had moise-traders arriving in a random fashion so that the market-maker could infer information from an imbalance of orders \citet{Kyle1985Continuous}.  However, though the idea that noise is a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, this exogenous construct may be an over-simplification that is not apprpropriate when looking for more understanding of speculation itself.  Institutional and psychological explanations for noise-trading suggest that there are endogenous elements that suggest non-random effects.The modelling of noise is related to the rise of rational expectations in economics.

@@

% Shliefer, Summers, DeLong etc identify the limits to arbitrage. The other writings are Battalio, Hatch and Jennings (1997), Harris and Schultz (1998) and Scheinkman and Xiong (2003).  Lowenstein and Willard (2006)

%It would be good to have some more modern ideas here.  Maybe they are just the empirical results. 
\subsection{Interplay between Informed and uninformed}
\citet[p. 1]{GortonNoise}  say that noise-traders are all those trading in markets for ``non-information-based reasons'', and that this was originally a solution to the problem of understanding or justifying the trading of securities: if economic agents know the fundamental value of a security how can some be buying and some be selling.  Will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of it. Informed traders, knowing that the current price is below fundamental value, can purchase from noise-traders, who either do not know or do not care.  Noise-traders can therefore create a gap between the price and the fundamtental value that will encourage informed traders to pay the cost of acquiring information. The divergence between this price and fundametnal value is essential to encourage te process of price discovery.

@@

 % Why do too extra  exist?  There are liquidity and behavioural explanations. There are also a number of models that show noise traders as momentum traders.

@@

One approach to this problem is to use the evidence that noise-traders utilise indicators or techniques that are believed to provide information when they do not.  \citet{ShillerFashion} discusses fashionable techniques.  In this way, speculation noise is associated with sentiment and measures of sentiment are then used as ways of measuring speculation. In addition,empirical work will often identify speculation or noise with momentum trading. However, this is controversial as from \citet{Jagadeesh} to the present, there is evidence that momentum trading is profitable citet(that latest paper in citulike).
